
% From https://www.overleaf.com/learn/latex/Glossaries

\makeglossaries % Prepare for adding glossary entries

\newglossaryentry{backup}
{
        name=backup,
        description={Copy of data that can be used to restore the original in case something happens to it.}
}

\newglossaryentry{exfiltration}
{
        name=data exfiltration,
        description={Transferring data out of an organization's network. Essentially data theft.}
}

\newglossaryentry{zero-day}
{
        name=zero-day,
        description={Computer security vulnerability that is unknown or unpatched.}
}

\newglossaryentry{malware}
{
        name=malware,
        description={Malicious software. General term for software that does damage to computers, IT systems, etc.}
}

\newglossaryentry{ransomware}
{
        name=ransomware,
        description={Type of malware that encrypts files, in order for the attacker to demand a ransom from their victims in exchange for the files being decrypted.}
}

\newglossaryentry{privilege escalation}
{
        name=privilege escalation,
        description={Attack where the attacker obtains unauthorized privileges/permissions, often by exploiting a vulnerability.}
}

\newglossaryentry{security controls}
{
        name=security controls,
        description={Noe her?}
}


\newglossaryentry{vm}
{
        name=VM,
        description={Short for Virtual Machine, a computer emulated in software.}
}

\newglossaryentry{OS}
{
        name=operating system,
        description={Software that manages hardware and provides services for users to interact with said hardware.}
}

\newglossaryentry{rdp}
{
        name=RDP,
        description={Remote Desktop Protocol. A Microsoft protocol for accessing computers via a graphical user interface remotely (over a network)}
}

\newglossaryentry{rsv}
{
        name=RSV,
        description={Recovery Services vault. A container for backup items and configurations in Azure}
}

\newglossaryentry{unmanaged}
{
        name=unmanaged service,
        description={When referring to cloud services, an unmanaged service is a service that is configured and maintained by the customer, and not by the cloud provider.}
}


% entries below are kept for reference

\newglossaryentry{bibliography}
{
        name=bibliography,
        plural=bibliographies,
        description={A list of the books referred to in a scholarly work,
typically printed as an appendix}
}

\newglossaryentry{maths}
{
    name=mathematics,
    description={Mathematics is what mathematicians do}
}


% --------------------
% ----- Acronyms -----
% --------------------

\newacronym{phd}{PhD}{philosophiae doctor}
\newacronym{CoPCSE}{CoPCSE@NTNU}{Community of Practice in Computer ScienceEducation at NTNU}
\newacronym{gcd}{GCD}{Greatest Common Divisor}
\newacronym{C&C}{C&C}{Command-and-control}
%\newacronym{rdp}{RDP}{Remote Desktop Protocol}
